Preparing for Potential Closure of European Airspaces due to Re-entering Space Objects

### 2025-05-02 paper.qmd

It is assumed the aircraft impact area to be much larger than the debris
object area so as to neglect the debris impact area.
The geometry of the aircraft is simplified and modelled as
a rectangular cuboid with dimensions of wingspan, length and height. 
Now, if the aircraft were stationary, the area exposed to a debris impact would
be the top view, while if the debris were stationary it would be the front view.
In reality an aircraft en-route is moving (mainly) horizontally and the debris 
vertically [^balistic_descent], which which allows to express the effective
exposed area, $\varepsilon_k$, as

@@

```{r collision}
#| out-width: "100%"
#| echo: false
knitr::include_graphics("media/example-image.png")
```

Collision risk probability. 
:::

@@

In few days (at the time of writing this paper) at 6h May 10, 2025
($\pm 2.8$ days), the KOSMOS 482 Descent Craft should reenter the
Earth's atmosphere [@kosmos482re-entry].
This is supposedly the lander module from a 1972 failed Soviet Venera mission
[@SpaceReviewKosmos] of \qty{480}{kg} weight and circa \qty{1}{m} diameter.
Being supposed to survive Venus's atmosphere which is much denser and hotter than Earth's atmosphere (surface pressure 93 times that of Earth and temperatures reaching \qty{467}{degreesCelsius} (\qty{872}{degreesFarenheit})).
So it could re-enter Earth intact and not break into smaller fragments some of which could burn up.

With what developed in this paper, we can derive the casuality risk and
determine the maximum value using data for the busiest day in 
the busiest 2024 week.

Once the re-entry track is known more or less accurately, the relevant
H3 cells can be

@@

<!-- future work ----------------------------------------------------------------------------->
Future work will explore the impact of megacostellations and the risk for
European airspace.

<!-- if any final ~ highly inspirational thought / words ------------------------------------->

# Appendix {.appendix}
# Orbital Elements {#sec-orbitals }

:::{#fig-orbital-elements layout="[40,-4,56]" fig-env="figure*" fig-pos="t"}

>>>>

We assume the aircraft impact area to be much larger than the debris
object area so as to neglet the debris impact area.
We can also simplify the geometry of the aircraft and model it as
a rectangular solid with dimesions of wingspan ($W$), length ($L$) and
height ($H$). 
Now, if the aircraft were stationary, the area exposed to a debris impact would
be the top view, while if the debris were stationary it would be the front view.
In reality an aircraft en-route is moving (mainly) horizontally and the debris 
vertically [^balistic_descent], which which allows to express the effective
exposed area, $\varepsilon_k$, as

@@

```{r collision}
#| out-width: "100%"
#| echo: false
knitr::include_graphics(here::here("media", "figures", "collision_expectation_2024-07-05_3.png"))
```

Averaged hourly collision expectation on 2024-07-05. 
:::

@@

In few days (at the time of writing of this paper) at 6h May 10, 2025
($\pm 2.8$ days), the KOSMOS 482 Descent Craft should reenter the
Earth's atmosphere [@kosmos482re-entry].
This is supposedly the lander module from a 1972 failed Soviet Venera mission
[@SpaceReviewKosmos] of $\qty{480}{kg}$ weight and approximately
$\qty{1}{m}$ diameter.
Being supposed to survive Venus's atmosphere which is much denser and hotter than Earth's atmosphere (surface pressure 93 times that of Earth and temperatures reaching \ang{467}C (\ang{872}F)),
it could re-enter Earth intact and not break into smaller fragments some of which could burn up.

Knowing that the debris inclination is $\ang{51.7}$,
see @fig-weighting-inclination-i, we can derive the maximum collision expectation using data for the busiest day in the busiest 2024 week as a
proxy for the traffic density on the day of reentry.

Once the reentry track is known, the H3 cells around \qty{35}{km} on each side 
can be identified to retrieve the values of collision expectancy.
It can then be decided whether it is advisable to close the airspace and
traffic around the area afftected.

@@

Future work will explore the impact of megacostellations and flight growth
forecast and what would happen in 20, 40, ..., 100 years in varios scenarios
of uncontrolled  reentries and the risk for European airspace.

# Appendix: Orbital Elements {#sec-orbitals}

:::{#fig-orbital-elements layout="[40,-4,56]" fig-env="figure*" fig-pos="h"}
